---
title: "ARIMA Models"
output: html_notebook
---

Non Seasonal ARIMA model
ARIMA means Auto Regressive Integrated Moving Average

xt is an ARIMAp,d,q model if ∆dxt is an ARMAp,q model

ARIMA(0,1,0) = random walk
ARIMA(0,1,1) without constant = simple exponential smoothing
ARIMA(0,2,1) without constant = linear exponential smoothing
ARIMA(1,1,2) with constant = damped-trend linear exponential smoothing


```{r}
library(forecast)
library(fpp2)
```



```{r}
autoplot(uschange[,"Consumption"]) +
xlab("Year") + ylab("Quarterly percentage change")
```

```{r}
Arima(uschange[,"Consumption"],order=c(2,0,2))
```

```{r}
auto.arima(uschange[,"Consumption"])
```

to obtain a good forecast and an understanding of the model
- let’s start by differencing the series if needed, in order to obtain something visually stationary
- look at the ACF and PACF plot ot identify possible models
- take eventually into account knowledge on the series (knwon autocorrelation. . . )
- estimate models and select the best one by AICc / AIC / BIC


```{r}
ggAcf(uschange[,"Consumption"])
```

```{r}
ggPacf(uschange[,"Consumption"])
```

May be an AR3 or an MA3

```{r}
Arima(uschange[,"Consumption"],order=c(3,0,0))
```

```{r}
Arima(uschange[,"Consumption"],order=c(0,0,3))
```

Predict
```{r}
fit=Arima(uschange[,"Consumption"],order=c(3,0,0))
autoplot(forecast(fit,h=10))
```

```{r}

```






