---
title: "DataSet Airpassengers from 1949 to 1960"
output: html_notebook
---

Libraries
```{r}
library(forecast)
library(ggplot2)
```

Loading dataset
```{r}
data(AirPassengers)
autoplot(AirPassengers)
```


```{r}
autoplot(decompose(AirPassengers,type="multiplicative"))+xlab('Year')
```

```{r}
length(AirPassengers)
```

remove seasonal pattern
```{r}
res1 = diff(AirPassengers,lag=12,differences=1)
autoplot(res1)
length(res1)
```

still a slight trend, apply diff again 
Rem: no need for lag=12, no seasonal pattern, allow to keep more points
```{r}
res2 = diff(res1,lag=1)
autoplot(res2)
length(res2)
```

check autocorrelation
acf of order 1 ~ -0.3 --> still correlation between xt and xt-1
```{r}
acf(res2)
```

Confirm with box test:
Check if white noise
p-value small (<0.05) --> HO rejected --> not white noise
```{r}
Box.test(res2,lag=10,type="Ljung-Box")$p.value
```

--> some autocorrelation to modelize.


********************************************************

Using Arima and heteroscedasticity

```{r}
data = AirPassengers

train = window(data,end=c(1958,12))
test = window(data,start=c(1959,1))

```

BoxCox transformation

```{r}
lambda = BoxCox.lambda(train)
l_train = BoxCox(train,lambda)
l_test = BoxCox(test,lambda)
autoplot(l_train,series='train') + autolayer(l_test,series='test')
```

```{r}
l_train %>% diff(lag=12) %>% diff() %>% ggtsdisplay()

```

```{r}
fit1=Arima(l_train, order=c(0,1,2), seasonal=c(0,1,1))
summary(fit1)
# fit1 %>% checkresiduals()
fit1 %>% residuals() %>% ggtsdisplay()
```

```{r}
autoplot(l_train) +
  autolayer(forecast(fit1,h=24),PI=FALSE,series='log arima') + 
  autolayer(l_test,series = 'log test') + 
  coord_cartesian(xlim = c(1958,1961))
```


```{r}
p = forecast(fit1,h=24)
forecast_inv = InvBoxCox(p$mean,lambda)

autoplot(train) +
  autolayer(test,series = 'Test') + 
  autolayer(forecast_inv,series = 'Sarima', PI=FALSE) + 
  coord_cartesian(xlim = c(1958,1961))


```




```{r}
fit2=Arima(train, order=c(0,1,2), seasonal=c(0,1,1),lambda = 'auto')
summary(fit2)
```

```{r}
autoplot(train) +
  autolayer(test,series = 'Test') + 
  autolayer(forecast(fit2,h=24),series = 'Sarima', PI=FALSE) + 
  coord_cartesian(xlim = c(1958,1961))
```


```{r}
fit3 = hw(train,h=24,lambda='auto')
```

```{r}
autoplot(train) +
  autolayer(test,series = 'Test') + 
  autolayer(forecast(fit2,h=24),series = 'Sarima', PI=FALSE) +
  autolayer(fit3$mean,series = 'HW') +
  coord_cartesian(xlim = c(1958, 1961))
```

```{r}
p = forecast(fit2,h=24)
print('Sarima RMSE')
print(sqrt(mean((p$mean-test)^2)))
print('HW RMSE')
print(sqrt(mean((fit3$mean-test)^2)))

```

