---
title: "Time Series Evaluation"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
--- 

```{r}
library(forecast)
library(ggplot2)
library(xlsx)
```


```{r}
rawdata = read.xlsx('Elec-train.xlsx',sheetIndex = 1)
names(rawdata)[2] = "Power"
names(rawdata)[3] = "Temp"

# dropping first 3 lines to simplify (will not have any effect on predictions...)
rawdata = tail(rawdata,n=4600)
data = ts(rawdata[,2:3],frequency = 96)
```


#### Data Exploration
```{r}
autoplot(data)
```

Looking at the 2 first periods
```{r}
autoplot(data)+
  coord_cartesian(xlim = c(1,3))
```

Power looks very noisy, maybe we can try to downsample the series to 1h instead of 15min for each observations
Holt winters was not working with freq = 96

```{r}
power_hour = c()
temp_hour = c()

for (i in 1:1150) {
  deb = 4*(i-1)+1
  fin = 4*(i-1)+4
  power_hour = c(power_hour,mean(data[deb:fin,"Power"]))
  temp_hour = c(temp_hour,mean(data[deb:fin,"Temp"]))
}

power_hour_ts = ts(power_hour,freq=24)
temp_hour_ts = ts(temp_hour,freq=24)

autoplot(data[,"Power"])+
  autolayer(power_hour_ts) +
  coord_cartesian(xlim = c(5,6))

autoplot(data[,"Temp"])+
  autolayer(temp_hour_ts) +
  coord_cartesian(xlim = c(5,10))
```

## Forecast without using Temperature
(use the last 24hours as test set to evaluate RMSE)

```{r}
train = window(power_hour_ts,end=c(46,22))
test = window(power_hour_ts,start=c(46,23),end=c(47,22))
```

```{r}
ggseasonplot(train)
```

#### Seasonal Holt Winters Additive (with and without Damping)

```{r}
fit1 = hw(train,h=24,seasonal = "additive",damped = FALSE)
fit2 = hw(train,h=24,seasonal = "additive",damped = TRUE)

print(paste("HW no damping RMSE : ",sqrt(mean((fit1$mean-test)^2))))
print(paste("HW with damping RMSE : ",sqrt(mean((fit2$mean-test)^2))))

```

#### Try using SARIMA

```{r}
train %>% diff(lag=24) %>% diff() %>% ggtsdisplay()
```

After diff for seaonality and diff to remove trend, 
it appears:
  - remaining peaks on pacf at 24,48,72... and on acf at 24 --> add seasonal MA

--> model model SARIMA (0,1,0)(0,1,1)

  - pacf exponential decrease and peak on acf at order 2 --> add non seasonal MA

--> try SARIMA (0,1,2)(0,1,1)

  - still some peak on acf at order 2 --> increase non seasonal MA
  
--> try SARIMA (0,1,4)(0,1,1)
  
```{r}
fit3 = Arima(train,order=c(0,1,4),seasonal = c(0,1,1))
fit3 %>% residuals() %>% ggtsdisplay()
```

Residuals look nice, check with box test

  --> high p-value : nothing to model anymore 

```{r}
fit3 %>% residuals() %>% Box.test()
```

Trying with auto.Arima for comparison

```{r}
fit4 = auto.arima(train)
```

Predict values and check RMSE

```{r}
pred3 = forecast(fit3,h=24)
pred4 = forecast(fit4,h=24)
print(paste("SARIMA (0,1,4)(0,1,1) : ",sqrt(mean((pred3$mean-test)^2))))
print(paste("Auto ARIMA : ",sqrt(mean((pred4$mean-test)^2))))
```

#### Try using AutoRegressive Neural Networks

With automatic parameters
```{r}
fit5 = nnetar(train)
print(fit5)
pred5 = forecast(fit5,h=24)
print(paste("NNET (25,1,13): ",sqrt(mean((pred5$mean-test)^2))))
```

Plot models versus Test

```{r}
autoplot(train) +
  autolayer(test,series="Test") + 
  autolayer(fit1,series="HW no Damping",PI=FALSE) +
  autolayer(fit2,series="HW with Damping",PI=FALSE) +
  autolayer(pred3,series="SARIMA (0,1,4)(0,1,1)",PI=FALSE) +
  autolayer(pred4,series="Auto Arima",PI=FALSE) +
  autolayer(pred5,series="NNET",PI=FALSE) +
  coord_cartesian(xlim = c(46.5,48))
```

### RECAP OF RMSE FOR PREDICTION WITHOUT TEMPERATURE

Best result is obtained with NNAR, 
slightly better than SARIMA(0,1,4)(0,1,1) but much more complicated.

```{r}
print(paste("HW no damping RMSE : ",sqrt(mean((fit1$mean-test)^2))))
print(paste("HW with damping RMSE : ",sqrt(mean((fit2$mean-test)^2))))
print(paste("SARIMA (0,1,4)(0,1,1) : ",sqrt(mean((pred3$mean-test)^2))))
print(paste("Auto ARIMA : ",sqrt(mean((pred4$mean-test)^2))))
print(paste("NNET (25,1,13): ",sqrt(mean((pred5$mean-test)^2))))
```

Use SARIMA for final predictions

```{r}
complete_power = window(power_hour_ts,end=c(46,22))
fitfinal = Arima(complete_power,order=c(0,1,4),seasonal = c(0,1,1))
predfinal = forecast(fitfinal,h=24)
autoplot(predfinal,PI=FALSE) + 
  coord_cartesian(xlim=c(46,48))
```

Extrapolate the final result

```{r}

pred_15 = c()
for (i in 1:24) {
  new_val = predfinal$mean[i]
  if (i != 24) {
    inc = (predfinal$mean[i+1]-predfinal$mean[i])/4
  }
  for (j in 1:4) {
    pred_15 = c(pred_15,new_val)
    new_val = new_val + inc
  }
}

pred_15_without_temp = as.data.frame(pred_15)
names(pred_15_without_temp)[1] = "Without_Temp"
```

## Forecast using Temperature
(use the last 24hours as test set to evaluate RMSE)

```{r}
train_temp = window(temp_hour_ts,end=c(46,22))
test_temp = window(temp_hour_ts,start=c(46,23),end=c(47,22))
```

Try to fit with Linear model with season and trend
```{r}
fit11=tslm(train~train_temp+season+trend)
CV(fit11)
checkresiduals(fit11,test='LB',plot=FALSE)
```

residuals not independant --> modelize them using dynamic regression model

```{r}
res11 = residuals(fit11)
res11 %>% diff(lag=24) %>% diff() %>% ggtsdisplay()
```

```{r}
fit12 = Arima(train, xreg = train_temp,order = c(0,1,3),seasonal = c(0,1,1))
fit12 %>% residuals() %>% ggtsdisplay()
```

Check if residuals are white noise
--> high pvalue, nothing more to model
```{r}
fit12 %>% residuals() %>% Box.test()
```

Compare with auto ARIMA

```{r}
fit13 = auto.arima(train,xreg = train_temp)
fit13 %>% residuals() %>% ggtsdisplay()
```

Predict values and check RMSE

```{r}
pred12 = forecast(fit12,xreg = test_temp, h=24)
pred13 = forecast(fit13,xreg = test_temp, h=24)
print(paste("SARIMA (0,1,3)(0,1,1) With TEMP: ",sqrt(mean((pred12$mean-test)^2))))
print(paste("Auto ARIMA With TEMP: ",sqrt(mean((pred13$mean-test)^2))))
```

Try With Neural NetWorks

```{r}
fit14 = nnetar(train, xreg = train_temp)
print(fit14)
pred14 = forecast(fit14,xreg = test_temp, h=24)
print(paste("NNET (25,1,14) With TEMP: ",sqrt(mean((pred14$mean-test)^2))))
```

```{r}
autoplot(train) +
  autolayer(test,series="Test") + 
  autolayer(pred12,series="SARIMA (0,1,3)(0,1,1) With TEMP",PI=FALSE) +
  autolayer(pred13,series="Auto Arima With TEMP",PI=FALSE) +
  autolayer(pred14,series="NNET With TEMP",PI=FALSE) +
  coord_cartesian(xlim = c(46.5,48))
```

### RECAP OF RMSE FOR PREDICTION WITH TEMPERATURE

Best result is obtained with NNAR, 
slightly better than SARIMA(0,1,3)(0,1,1) but much more complicated.

```{r}
print(paste("SARIMA (0,1,3)(0,1,1) With TEMP: ",sqrt(mean((pred12$mean-test)^2))))
print(paste("Auto ARIMA With TEMP: ",sqrt(mean((pred13$mean-test)^2))))
print(paste("NNET (25,1,14) With TEMP: ",sqrt(mean((pred14$mean-test)^2))))
```

Use SARIMA for final predictions

```{r}
complete_power = window(power_hour_ts,end=c(46,22))
complete_temp = window(temp_hour_ts,end=c(46,22))
fitfinal_temp = Arima(complete_power,xreg = complete_temp, order=c(0,1,3),seasonal = c(0,1,1))
predfinal_temp = forecast(fitfinal_temp, xreg = complete_temp, h=24)
autoplot(predfinal_temp,PI=FALSE) + 
  coord_cartesian(xlim=c(46,48))
```

Extrapolate the final result 

```{r}

pred_15 = c()
for (i in 1:24) {
  new_val = predfinal_temp$mean[i]
  if (i != 24) {
    inc = (predfinal_temp$mean[i+1]-predfinal_temp$mean[i])/4
  }
  for (j in 1:4) {
    pred_15 = c(pred_15,new_val)
    new_val = new_val + inc
  }
}

pred_15_with_temp = as.data.frame(pred_15)
names(pred_15_with_temp)[1] = "With_Temp"
```


## Write Final preidctions to xls

```{r}
pred_15 = cbind(pred_15_without_temp,pred_15_with_temp)

write.xlsx(pred_15,'A-Jourdan.xlsx',col.names = TRUE,row.names = FALSE)
```

## Compare With and Without temperature models
slightly better with temperature effect.

```{r}
print(paste("SARIMA (0,1,4)(0,1,1) : ",sqrt(mean((pred3$mean-test)^2))))
print(paste("SARIMA (0,1,3)(0,1,1) With TEMP: ",sqrt(mean((pred12$mean-test)^2))))
```

```{r}
autoplot(data) +
  autolayer(ts(pred_15_without_temp,freq=96,start = c(47,89)),series="Without TEMP") +
  autolayer(ts(pred_15_with_temp,freq=96,start = c(47,89)),series="With TEMP") +
  coord_cartesian(xlim = c(47,49),ylim = c(140,330))
```

